An algorithm for finding the global extremum of a partially defined function\thanks{This work was supported by the Russian Science Foundation, project No. 21-11-00204.}

### 2023-08-28 2023/MMST/samplepaper.tex
minor corrections

\begin{figure}
\includegraphics[width=\textwidth]{fig1.png}
\caption{Examples of the distributions of the search trials in the non-computability regions in the first approach to the imputation of the values} \label{fig1}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{fig2.png}
\caption{Examples of the distributions of the search trials in the non-computability regions in the second approach to the imputation of the values} \label{fig2}
\end{figure}

This approach resolves partly the problems of the first approach outlined above. However, it doesn’t solve the problem of excess trials and requires a considerable increasing of the parameter $r$ for more complex test problem. The examples for the one-dimensional problems using the first two approaches are presented in ~\ref{fig1} and ~\ref{fig2} where the trials falling into the region with the undefined values are marked by the black points and the computable points of the search domain – by the green ones.

\subsubsection{The third approach.} To restore the objective function values in the points of the intervals with two non-computable points as the averages of the neighbor points and to apply corresponding computation rule from formula (\ref{eq9}) assuming the points to be computable ones. If the neighbor points are non-computable as well, to use formula (\ref{eq14}). In the case of an interval with one non-computable point, to apply formulae (\ref{eq15}) and (\ref{eq16}).
\begin{figure}
\includegraphics[width=\textwidth]{fig3.png}
\caption{Imputation of the non-computable values by the averages of the neighbor points. Approach \#3} \label{fig3}
\end{figure}

Essential drawbacks of this approach are not utilizing the information on the coordinates aw well as seldom executing the computation rule added because of searching the substituting values among the nearest neighbors only.

\subsubsection{The fourth approach.} To restore the objective function values in the points for the intervals with two non-computable points as the averages of the nearest computable points taking into account the location coordinates i. e.
\begin{equation}\label{eq17} 
\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \| x-x' \|}{\| x''-x' \|},
\end{equation}
where $z'$ and $z''$ are the objective function values in the computable points $x'$ and $x''$  nearest to the interval.
The rest of the approach is identical to the third one. Formula (\ref{eq14}) in this case will be applied when the non-computable region is located near the boundary point only.
\begin{figure}
\includegraphics[width=\textwidth]{fig4.png}
\caption{Imputation of non-computable values by averages of the neighboring points. Approach \#4} \label{fig4}
\end{figure}
\begin{figure}
\includegraphics[width=\textwidth]{fig5.png}
\caption{Imputation of non-computable values by averages of the minorant and the majorant. Approach \#5} \label{fig5}
\end{figure}

\subsubsection{The fifth approach.} To utilize the search information and the suggestions on the function form most completely, one can restore the objective function values in the points taking onto account the H{\"o}lder condition.
\paragraph{Step 1} Find the end points of the crossing of majorant and minorant
\begin{equation}\label{eq18} 
\hat{x}',\hat{x}''=\frac {x'+x''}{2}\pm \text{sign}(z''-z')\cdot \frac {1}{2r} \cdot {\left(\frac {z''-z'}{\mu}\right)}^N,
\end{equation}
\paragraph{Step 2} Restore the objective function values according to the rule
\begin{equation}\label{eq19} 
\tilde{z}(x)=
  \begin{cases}
    z', & {\quad x \leq \hat{x}',}\\
    \frac {z'+z''}{2}- \frac {r \mu}{2} (\| x - x' \| + \| x - x'' \|), & {\quad \hat{x}' < x < \hat{x}'',}\\
    z'',  & {\quad x \geq \hat{x}''.}
  \end{cases}
\end{equation}
This rule can be interpreted as an average value of the minorant and the majorant for a function satisfying the H{\"o}lder condition. Let us give the description of the modification of the global search algorithm in the case of not everywhere computable function based on this approach.

\subsubsection{Description of the modification} The algorithm remains the same with an introduction of the following minor additions in the computing formulae in the case of falling into the non-computable search regions:
\begin{enumerate} 
  \item The index $v_i=v(x_i)$ is determined by the formula (\ref{eq12}).
  \item When computing current lower estimate of the H{\"o}lder constant $K$ in formula (\ref{eq7}), use the intervals with two computable points only.
  \item The best current objective function value is determined for different values of the index
\begin{equation}\label{eq20} 
z_v^*=
  \begin{cases}
    -\varepsilon _r, & {\quad v=0 ,}\\
    \min \left\{ \phi (x_i): v(x_i)=1, 1 \leq i \leq k \right\}, & {\quad v=1 ,}
  \end{cases}
\end{equation}
where $\varepsilon _r$ is a parameter of the method.

>>>>

\begin{figure}
\includegraphics[width=\textwidth]{fig1.png}
\caption{Examples of the search trials distribution in the non-computability regions for the first approach to the values imputation} \label{fig1}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{fig2.png}
\caption{Examples of the search trials distribution in the non-computability regions for the second approach to the values imputation} \label{fig2}
\end{figure}

This approach partly resolves the problems of the first approach outlined above. However, it doesn’t solve the problem of excess trials and requires a considerable increasing of the parameter $r$ for more complex test problem. The examples for the one-dimensional problems using the first two approaches are presented in Fig.~\ref{fig1} and Fig.~\ref{fig2} where the trials falling into the region with the undefined values are marked by the black points and the computable points of the search domain are marked by the green ones.

\subsubsection{The third approach.} To restore the objective function values at the points of the intervals with two non-computable points as the averages of the neighbor points and to apply corresponding computation rule from formula (\ref{eq9}) assuming the points to be computable ones. If the neighbor points are non-computable as well, use formula (\ref{eq14}). In the case of an interval with one non-computable point, apply formulae (\ref{eq15}) and (\ref{eq16}).
\begin{figure}
\includegraphics[width=\textwidth]{fig3.png}
\caption{Imputation of the non-computable values by the averages of the neighbor points. Approach \#3} \label{fig3}
\end{figure}

Essential drawbacks of this approach are not utilizing the information on the coordinates aw well as seldom executing the computation rule added because of searching the substituting values only among the nearest neighbors.

\subsubsection{The fourth approach.} To restore the objective function values in the points for the intervals with two non-computable points as the averages of the nearest computable points taking into account the coordinates, i.e.
\begin{equation}\label{eq17} 
\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \| x-x' \|}{\| x''-x' \|},
\end{equation}
where $z'$ and $z''$ are the objective function values in the computable points $x'$ and $x''$  nearest to the interval.
The rest of the approach is identical to the third one. Formula (\ref{eq14}) in this case will be applied when the non-computable region is located only near the boundary point.
\begin{figure}
\includegraphics[width=\textwidth]{fig4.png}
\caption{Imputation of non-computable values by averages of the neighboring points. Approach \#4} \label{fig4}
\end{figure}
\begin{figure}
\includegraphics[width=\textwidth]{fig5.png}
\caption{Imputation of non-computable values by averages of the minorant and the majorant. Approach \#5} \label{fig5}
\end{figure}

\subsubsection{The fifth approach.} To utilize the search information and the suggestions on the function form most completely, one can restore the objective function values in the points taking onto account the H{\"o}lder condition.
\paragraph{Step 1} Find the end points of the crossing of majorant and minorant
\begin{equation}\label{eq18} 
\hat{x}',\hat{x}''=\frac {x'+x''}{2}\pm \text{sign}(z''-z')\cdot \frac {1}{2r} \cdot {\left(\frac {z''-z'}{\mu}\right)}^N,
\end{equation}
\paragraph{Step 2} Restore the objective function values according to the rule
\begin{equation}\label{eq19} 
\tilde{z}(x)=
  \begin{cases}
    z', & {\quad x \leq \hat{x}',}\\
    \frac {z'+z''}{2}- \frac {r \mu}{2} (\| x - x' \| + \| x - x'' \|), & {\quad \hat{x}' < x < \hat{x}'',}\\
    z'',  & {\quad x \geq \hat{x}''.}
  \end{cases}
\end{equation}
The value $\tilde{z}(x)$ can be interpreted as an average value of the minorant and the majorant for a function satisfying the H{\"o}lder condition. Let us give the description of the modification of the global search algorithm in the case of partially defined function based on this approach.

\subsubsection{Description of the modification.} The algorithm remains the same with an introduction of the following minor additions in the computing formulae in the case of falling into the non-computable search regions.
\begin{enumerate} 
  \item The index $v_i=v(x_i)$ is determined by the formula (\ref{eq12}).
  \item When computing current lower estimate of the H{\"o}lder constant $K$ in formula (\ref{eq7}), use only the intervals with two computable boundary points.
  \item The value $z_v^*$ depends on the index value
\begin{equation}\label{eq20} 
z_v^*=
  \begin{cases}
    -\varepsilon _r, & {\quad v=0 ,}\\
    \min \left\{ \phi (x_i): v(x_i)=1, 1 \leq i \leq k \right\}, & {\quad v=1 ,}
  \end{cases}
\end{equation}
where $\varepsilon _r > 0$ is a parameter of the method.
